Sketch Recognition using Particle Swarm Optimizations

### 2009-05-07 trunk/paper2/Paper2.tex
saving before chainging the abstract6

 \begin{figure}
	\centering		
	 \includegraphics[scale=0.3]{images/swarmAgentTest.jpg.eps}
	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
	 	\label{fig:swarmtesting}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

\begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{images/swarmParamterC.jpg.eps}
	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
	 	\label{fig:swarmtesting2}
	%\caption{Experiments results :} a)    b) c)

@@

\begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{images/LDAlg.jpg}
	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
	 	\label{fig:testLD}
	%\caption{Experiments results :} a)    b) c) 
\end{figure} 
\subsection{Shape Complexity Analysis}
\label{sec:ShapeComplexityExperiments}

@@

\begin{figure*}
	\centering
		\includegraphics[scale=0.5]{images/LDsymbols.jpg}
	\caption{Symbols Comparison on LD-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:LDtest2}
\end{figure*}  
The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} and \ref{fig:LDtest2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}). It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols. The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines. This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself. Figure \ref{fig:LDtest2} shows that in logic design symbols \textsl{AlgS2} achieved better performance than \textsl{Alg3} \textsl{AlgS1} this emphases that complex shapes that are combination of lines and arc are better segmented using \textsl{AlgS2}. Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result. This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation. Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols. The penalty for this improved performance is the computational time required to run both swarm algorithms. Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols. The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles. This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them. This lead us to the next experiment to choose best set of feature to recognize symbols. 
\begin{table*}
	\centering
	%\small
%	\scalebox{0.7}{
		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
 \include{conf}
 		%\end{tabular}
 	%}
	\caption{Confusion Matrix of Hs-DB}
	\label{tab:ConfusionMatrix}
	%\end{minipage}
\end{table*}
% \begin{table*}
% 	\centering
% 	%\small
%	\scalebox{0.7}{
% 		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
%  \include{conf2}
%  		%\end{tabular}
%  	%}
% 	\caption{Confusion Matrix of DL-DB}
% 	\label{tab:ConfusionMatrix}
% 	%\end{minipage}
% \end{table*}
\subsection{Features Analysis}
\label{sec:featexp}
Different feature sets are tested to determine the best features that can used in sketch and symbol recognition. Figure\ref{fig:testFeaturesAll} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}). Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features. This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}. Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.
 \begin{figure*}
	\centering
		\includegraphics[scale=0.5]{images/featuresdifference.jpg}
	\caption{Feature Comparison} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeaturesAll}
\end{figure*}  
 We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous. It is only based on what the user intends to draw while sketching the symbol. 
 The main features affected by the segmentation stage are the geometric features computed in\textbf{FS1 }. Hence those are the features we used to estimate the efficiency of the segmentation algorithm. The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} . The figure shows that \textsl{AlgS2} achieve better results.  %effect of the segmentation algorithm 
 %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block. The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}). Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features. 
\begin{figure}
	\centering
		\includegraphics[scale=0.4]{images/featureFS1only.jpg}
	\caption{} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeatonly}

>>>>

 \begin{figure}
	\centering		
	 \includegraphics[scale=0.3]{images/swarmAgentTest.jpg}
	 	\caption{Swarm testing} Segmentation Error Vs. Number of agents
	 	\label{fig:swarmtesting}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

\begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{images/swarmParamterC.jpg}
	 	\caption{Swarm testing} Segmentation Error Vs. PSO parameters.
	 	\label{fig:swarmtesting2}
	%\caption{Experiments results :} a)    b) c)

@@

\begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{images/LDAlg.jpg}
	 	\caption{Algorithm comparison on LD-DB} The recognition rate of different algorithms. 
	 	\label{fig:testLD}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

\begin{figure}
	\centering		
	 \includegraphics[scale=0.4]{images/AlgEL.jpg}
	 	\caption{Algorithm comparison on EL-DB} The recognition rate of different algorithms. 
	 	\label{fig:testEL}
	%\caption{Experiments results :} a)    b) c) 
\end{figure}

\subsection{Shape Complexity Analysis}
\label{sec:ShapeComplexityExperiments}

@@

\begin{figure*}
	\centering
		\includegraphics[scale=0.5]{images/LDsymbols.jpg}
	\caption{Symbols Comparison on LD-DB symbols}.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:LDtest2}
\end{figure*}  
The second experiment we implemented was to investigate the effect of symbol complexity and type on the recognition rate. Figure \ref{fig:test2} and \ref{fig:LDtest2} shows the achieved accuracy of each symbol by our algorithms compared to \cite{earlyprocess} (\textsl{Alg3}). It is clearly noted that symbols that have only line segments achieve higher accuracy rate than other symbols. The results indicate that algorithm \textsl{AlgS1} achieve better performance than algorithm \textsl{AlgS2} in the symbols that consist only of lines. This is understandable because algorithm \textsl{AlgS1} divides strokes into line segments only but \textsl{AlgS2} is able to divide strokes into lines and curves based on the minimum error of the segment itself. Figure \ref{fig:LDtest2} shows that in logic design symbols \textsl{AlgS2} achieved better performance than \textsl{Alg3} \textsl{AlgS1} this emphases that complex shapes that are combination of lines and arc are better segmented using \textsl{AlgS2}. Algorithm \textsl{Alg3} gives good performance as long as the symbols consist of combination of lines and curves, if the stroke consists of only curve or lines the algorithm may lead to wrong segmentation result. This is because the system divides the stroke first to line segments then tries to decide if each segment can be represent better as a curve unlike algorithm \textsl{AlgS2} where the curve segments are tested while choosing the best segmentation. Combining both algorithm \textsl{AlgS1} and \textsl{AlgS2} improved the recognition rate of all symbols. The penalty for this improved performance is the computational time required to run both swarm algorithms. Table \ref{tab:ConfusionMatrix} shows the confusion matrix of symbols. The table shows that errors are only between two sets of symbols moon with triangles and trapezoid with triangles. This observation is understandable as the symbols are visually similar but the system must be able to differentiate between them. This lead us to the next experiment to choose best set of feature to recognize symbols. 
\begin{table*}
	\centering
	%\small
%	\scalebox{0.7}{
		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
 \include{conf}
 		%\end{tabular}
 	%}
	\caption{Confusion Matrix of Hs-DB}
	\label{tab:ConfusionMatrix}
	%\end{minipage}
\end{table*}
% \begin{table*}
% 	\centering
% 	%\small
%	\scalebox{0.7}{
% 		%	\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}\hline 
%  \include{conf2}
%  		%\end{tabular}
%  	%}
% 	\caption{Confusion Matrix of DL-DB}
% 	\label{tab:ConfusionMatrix}
% 	%\end{minipage}
% \end{table*}
\subsection{Features Analysis}
\label{sec:featexp}
Different feature sets are tested to determine the best features that can used in sketch and symbol recognition. Figure\ref{fig:testFeaturesAll} shows the result of different feature sets from the basic four sets (\textbf{FS1,FS2,FS3,FS4}). Result shows that (\textbf{FS2}) Rubine features \cite{gestureexample12} gives the worst results when used alone without any other features. This is because they are mainly computed for single stroke gestures and fare bad in multi-stroke symbols \cite{compareFeaturSVM}. Features \textbf{(FS4)} gives good results but it is improved by adding structural features \textbf{(FS1)}.
 \begin{figure*}
	\centering
		\includegraphics[scale=0.5]{images/featuresdifference.jpg}
	\caption{Feature Comparison on Hs-DB accuracy} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeaturesAll}
\end{figure*}

 \begin{figure*}
	\centering
		\includegraphics[scale=0.34]{images/featELc.jpg}
	\caption{Feature Comparison on EL-DB} The effect of different features on accuracy.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:ELtestFeaturesAll}
\end{figure*}  
 We could not test the result of the segmentation algorithm directly due to the fact that the correct segmentation is highly ambiguous. It is only based on what the user intends to draw while sketching the symbol. 
 The main features affected by the segmentation stage are the geometric features computed in \textbf{FS1}. Hence those are the features we used to estimate the efficiency of the segmentation algorithm. The recognition accuracy using only the geometrical features is reported in Figure \ref{fig:testFeatonly} . The figure shows that \textsl{AlgS2} achieve better results.  %effect of the segmentation algorithm 
 %Figure \ref{fig:SampleSeg} shows a sample of the output of the segmentation block. The result of segmentation block is tested by their effect on the final recognition results using the features that will be highly affected using different algorithms like \textbf{FS1} and \textbf{FS4} (Section\ref{sec:Recognition}). Figure \ref{fig:testFeatonly} shows the result of the different algorithms using only \textbf{FS1 }features. 
\begin{figure}
	\centering
		\includegraphics[scale=0.4]{images/featureFS1only.jpg}
	\caption{The effect of segmentation on Hs-DB accuracy} %The effect of symbol complexity.  %The graph shows the recognition rate of each symbol using different algorithms. 
	\label{fig:testFeatonly}
\end{figure}
